# Solutions: Using configuration files in your RAP pipeline

This document provides step-by-step solutions for adding and using configuration parameters in a RAP pipeline. Each step matches the corresponding exercise notebook and demonstrates a modular approach.

```{r}
# Setup code for solutions

library(yaml)
source(here::here("R", "cleaning.R"))
source(here::here("R", "io.R"))
source(here::here("R", "processing.R"))
source(here::here("R", "report.R"))
source(here::here("R", "utils.R"))

# Construct the path to the YAML configuration file
config_path <- file.path("config/user_config.yaml")

# Load the YAML configuration file
config <- yaml::read_yaml(config_path)

input_path = config$input_path

# I/O Read data
health_df <- read_health_data(input_path)

# Output path
output_path <- "data/outputs/cleaned/health_data_cleaned"
```

## Step 1 Solution: Add a new parameter to the config file

Open `config/user_config.yaml` and add the following parameter:

```         
min_height_cm: 155  # Minimum height (cm) to include in analysis
```

This parameter allows you to control which rows are included based on height. Save the file after editing.

## Step 2 Solution: Read config parameters in a modular way

Create a function to read config values in a new module, e.g., `R/io.R`:

```{r}
# R/main.R

min_height <- config$min_height
```

```{r}
print(min_height)
```

## Step 3 Solution: Use config parameters in your pipeline

Update your pipeline (e.g., in `main.R` and the `cleaning.R` module) to use the config-driven parameter for filtering:

```{r}
filter_by_min_height <- function(df, min_height) {
  return(df[df$height_cm >= min_height, ])
}
```

Example usage in main.R

```{r}
source(here::here("R", "cleaning.R"))
source(here::here("R", "io.R"))

min_height <- config$min_height

filtered_df <- filter_by_min_height(health_df, min_height)
```

## Step 4 Solution: Add and use another config parameter (output format)

Add another parameter to your config file:

```         
output_format: csv  # Output file format: csv or xlsx
```

Create a modular function to save output in the desired format, e.g., in `R/io.R`:

```{r}
library(openxlsx)
save_output <- function(df, output_path, output_format) {
  if (output_format == "csv") {
    write.csv(df, paste0(output_path, ".csv"), row.names = FALSE)
  } else if (output_format == "xlsx") {
    write.xlsx(df, paste0(output_path, ".xlsx"), row.names = FALSE)
  } else {
    stop("Unsupported output format.")
  }  
}
```

```{r}
source(here::here("R", "io.R"))

output_format <- config$output_format
save_output(filtered_df, output_path, output_format)
```

