# Solutions: Unit Testing Exercises

This notebook provides step-by-step solutions for writing and running unit tests in your RAP pipeline using testthat. Each solution matches the corresponding exercise notebook and is designed for beginners.

## Solution 1: Write a simple unit test for a new function

Here are example unit tests for the `flag_missing` and `impute_by_group` functions in `R/cleaning.R`:
```{r}
library(testthat)
source("R/cleaning.R")
testthat::test_that("test flag_missing", {
  df <- data.frame(
    height_cm = c(170, NA, 160),
    weight_kg = c(70, 80, NA))
  expected_output <- data.frame(
    height_cm = c(170, NA, 160),
    weight_kg = c(70, 80, NA),    
    height_cm_imputed = c(FALSE, TRUE, FALSE),
    weight_kg_imputed = c(FALSE, FALSE, TRUE))
  
  flagged = flag_missing(df)
  testthat::expect_equal(as.data.frame(flagged), expected_output)
})
```

```{r}
library(testthat)
source("R/cleaning.R")
testthat::test_that("test impute_by_group", {
  df <- data.frame(
    height_cm = c(170, NA, 160, 150, 190, 180, NA, NA, 170, NA),
    weight_kg = c(70, 80, 60, 50, 90, 70, 80, 100, 50, 90),
    sex = c("M", "M", "F", "F", "M", "M", "F", "M", "F", "M"))
  expected_output <- c(170, 180, 160, 150, 190, 180, 160, 180, 170, 180)
  output <- impute_by_group(df, "height_cm", "sex")
  testthat::expect_equal(round(output, 1), round(expected_output, 1))
})
```

## Solution 2: Run your unit tests

**Task:**
- To run tests in the `tests/` folder you must open each test one at a time and click the `run tests` button in the top right corner of the source pane.
- Check the output to see if your tests pass or fail.
 
**Expected output:**
- All tests should pass. If a test fails, check the error message and fix your code or tests.

## Solution 3: Stretch - Check test coverage

Run the following commands:
```{r}
library(covr)
 cov <- covr::file_coverage("R/utils.R", "tests/test_utils.R")
 zero_lines <- covr::zero_coverage(cov)
 print(zero_lines)
# Note: The first argument is the path to the function, the second argument is the path to the test function
```
**Expected output:**
- You will see a report showing the percentage of code covered by tests. Aim for high coverage, but focus on testing important logic.


## Solution 4: Stretch - Try parameterisation in testthat

Here are examples using `with_parameters_test_that` for `flag_missing` and `impute_by_group`. Parameterisation lets you run the same test with different inputs, making your tests more robust and easier to maintain.

```{r}
library(testthat)
library(patrick)

patrick::with_parameters_test_that("test flag_missing", {
    flagged <- flag_missing(df, columns)
    testthat::expect_equal(flagged$height_cm_imputed, expected_height_flags)
    testthat::expect_equal(flagged$weight_kg_imputed, expected_weight_flags)
  },
  df = list(
    data.frame(height_cm = c(170, NA), weight_kg = c(70, NA)),
    data.frame(height_cm = c(NA, NA), weight_kg = c(NA, 80))
  ),
  columns = list(
    c("height_cm", "weight_kg"),
    c("height_cm", "weight_kg")
  ),
  expected_height_flags = list(
    c(FALSE, TRUE),
    c(TRUE, TRUE)
  ),
  expected_weight_flags = list(
    c(FALSE, TRUE),
    c(TRUE, FALSE)
  )
)
```

```{r}
library(testthat)
library(patrick)
library(covr)
library(dplyr)
impute_by_group <- function(df, col, group_col) {
  group_means <- df %>%
  dplyr::group_by(!!rlang::sym(group_col)) %>%
  dplyr::mutate(group_mean = mean(!!rlang::sym(col), na.rm = TRUE)) %>%
  dplyr::pull(group_mean)
  result <- ifelse(is.na(df[[col]]), group_means, df[[col]])
  result <- ifelse(is.na(result), mean(df[[col]], na.rm = TRUE), result)
  result <- round(result, 1)
  return(result)
}

patrick::with_parameters_test_that("test impute_by_group", {
  imputed <- impute_by_group(df, col, group_col)
  testthat::expect_equal(imputed, expected)
  },
  df = list(
    data.frame(height_cm = c(170, NA, 160), sex = c("M", "F", "F")),
    data.frame(height_cm = c(NA, NA, 150), sex = c("M", "M", "F"))
  ),
  col = c("height_cm", "height_cm"),
  group_col = c("sex", "sex"),
    expected = list(
    height_cm = c(170, 160, 160),
    height_cm = c(150, 150, 150)
  )
)
```

