# Solutions: Unit Testing Exercises

This notebook provides step-by-step solutions for writing and running unit tests in your RAP pipeline using testthat. Each solution matches the corresponding exercise notebook and is designed for beginners.

## Exercise 1 Solution: Review and Adapt an existing unit

Open `tests/testthat/test_cleaning.R` and run the test.

After the test passes, change the DataFrame to the following DataFrame and run the test again:

```{r}
df <- data.frame(
  height_cm = c(170, NA, 160, 165),
  weight_kg = c(70, 80, 75, 68),
  diagnosis = c(NA, NA, "A", "B"),
  smoker = c(NA, "No","Yes", NA),
  sex = c("m", "f", "M", "F"),
  stringsAsFactors = FALSE
)
```

Notice how the check for the missing smoker value fails. The test checks column 0 for a "smoker" value of "No", with the new DataFrame the first 2 columns are imputed due to missing diagnosis leaving the DataFrame looking like this:

```{r}
df <- data.frame(
  height_cm = c(160, 165),
  weight_kg = c(75, 68),
  diagnosis = c("A", "B"),
  smoker = c("Yes", NA),
  sex = c("M", "F"),
  stringsAsFactors = FALSE
)
```

The test checks column 0 and expects the value "No" in the "smoker" column, instead it receives the value "Yes" and therefore the test fails.
In order for the test to pass either change the column that the test checks or change the expected "smoker" value to "Yes" instead of "No".

The original expect statement looks like this:

```{r}
testthat::expect_equal(cleaned$smoker[1], "No")
```

The changed expect statement should look like either:

```{r}
testthat::expect_equal(cleaned$smoker[1], "Yes")
```

or

```{r}
testthat::expect_equal(cleaned$smoker[2], "No")
```

## Exercise 2 Solution: Write a simple unit test for a new function

Here are example unit tests for the `flag_missing` and `impute_by_group` functions in `R/cleaning.R`:
```{r}
library(testthat)
source(here::here("R", "cleaning.R"))
testthat::test_that("test flag_missing", {
  df <- data.frame(
    height_cm = c(170, NA, 160),
    weight_kg = c(70, 80, NA))
  expected_output <- data.frame(
    height_cm = c(170, NA, 160),
    weight_kg = c(70, 80, NA),    
    height_cm_imputed = c(FALSE, TRUE, FALSE),
    weight_kg_imputed = c(FALSE, FALSE, TRUE))
  
  flagged = flag_missing(df)
  testthat::expect_equal(as.data.frame(flagged), expected_output)
})
```

```{r}
library(testthat)
source(here::here("R", "cleaning.R"))
testthat::test_that("test impute_by_group", {
  df <- data.frame(
    height_cm = c(170, NA, 160, 150, 190, 180, NA, NA, 170, NA),
    weight_kg = c(70, 80, 60, 50, 90, 70, 80, 100, 50, 90),
    sex = c("M", "M", "F", "F", "M", "M", "F", "M", "F", "M"))
  expected_output <- c(170, 180, 160, 150, 190, 180, 160, 180, 170, 180)
  output <- impute_by_group(df, "height_cm", "sex")
  testthat::expect_equal(round(output, 1), round(expected_output, 1))
})
```

## Exercise 3 Solution: Run your unit tests

**Task:**
- To run tests in the `tests/` folder you must open each test one at a time and click the `run tests` button in the top right corner of the source pane.
- Check the output to see if your tests pass or fail.
 
**Expected output:**
- All tests should pass. If a test fails, check the error message and fix your code or tests.

## Exercise 4 Solution: Stretch - Check test coverage

Run the following commands:
```{r}
library(covr)
 cov <- covr::file_coverage("R/utils.R", "tests/test_utils.R")
 zero_lines <- covr::zero_coverage(cov)
 print(zero_lines)
# Note: The first argument is the path to the function, the second argument is the path to the test function
```
**Expected output:**
- You will see a report showing the percentage of code covered by tests. Aim for high coverage, but focus on testing important logic.


## Exercise 5 Solution: Stretch - Try parameterisation in testthat

Here are examples using `with_parameters_test_that` for `flag_missing` and `impute_by_group`. Parameterisation lets you run the same test with different inputs, making your tests more robust and easier to maintain.

```{r}
library(testthat)
library(patrick)

patrick::with_parameters_test_that("test flag_missing", {
    flagged <- flag_missing(df, columns)
    testthat::expect_equal(flagged$height_cm_imputed, expected_height_flags)
    testthat::expect_equal(flagged$weight_kg_imputed, expected_weight_flags)
  },
  df = list(
    data.frame(height_cm = c(170, NA), weight_kg = c(70, NA)),
    data.frame(height_cm = c(NA, NA), weight_kg = c(NA, 80))
  ),
  columns = list(
    c("height_cm", "weight_kg"),
    c("height_cm", "weight_kg")
  ),
  expected_height_flags = list(
    c(FALSE, TRUE),
    c(TRUE, TRUE)
  ),
  expected_weight_flags = list(
    c(FALSE, TRUE),
    c(TRUE, FALSE)
  )
)
```

```{r}
library(testthat)
library(patrick)
library(covr)
library(dplyr)
impute_by_group <- function(df, col, group_col) {
  group_means <- df %>%
  dplyr::group_by(!!rlang::sym(group_col)) %>%
  dplyr::mutate(group_mean = mean(!!rlang::sym(col), na.rm = TRUE)) %>%
  dplyr::pull(group_mean)
  result <- ifelse(is.na(df[[col]]), group_means, df[[col]])
  result <- ifelse(is.na(result), mean(df[[col]], na.rm = TRUE), result)
  result <- round(result, 1)
  return(result)
}

patrick::with_parameters_test_that("test impute_by_group", {
  imputed <- impute_by_group(df, col, group_col)
  testthat::expect_equal(imputed, expected)
  },
  df = list(
    data.frame(height_cm = c(170, NA, 160), sex = c("M", "F", "F")),
    data.frame(height_cm = c(NA, NA, 150), sex = c("M", "M", "F"))
  ),
  col = c("height_cm", "height_cm"),
  group_col = c("sex", "sex"),
    expected = list(
    height_cm = c(170, 160, 160),
    height_cm = c(150, 150, 150)
  )
)
```

