# Exercise: using packages and functions

In this exercise, you'll practice refactoring a single, monolithic script into a
modular pipeline using functions and separate files. This is a key RAP best
practice for reproducibility and maintainability.

**Learning objectives:**
- Understand the difference between monolithic and modular code
- Practice breaking up a large script into reusable functions
- Learn how to organise code into modules for analysis and reporting
- Learn how to handle missing data using imputation

```{r setup, include = FALSE}
# Setup code for exercises

library(dplyr)

input_path <- here::here("data", "inputs", "health_data.csv")

# Load data
df <- read.csv(input_path)

# print the first 5 rows of health data
head(df, 5)
```

## Exercise 1: Review the Monolithic Script

Below is a single script that identifies missing values and performs imputation.
Your task is to refactor this into modular code.

**Tasks:**
- Convert the below into functions
- Extend the function to impute missing height and weight using group means
(e.g., by gender or diagnosis)
- Add a new column to flag which values were imputed

```{r}
# Impute missing height and weight with column means
# Flag missing values
df$height_cm_imputed <- is.na(df$height_cm)
df$weight_kg_imputed <- is.na(df$weight_kg)

df <- dplyr::group_by(df, gender)
df <- dplyr::mutate(df,
  height_cm = ifelse(is.na(height_cm), mean(height_cm, na.rm = TRUE), height_cm),
  weight_kg = ifelse(is.na(weight_kg), mean(weight_kg, na.rm = TRUE), weight_kg)
)
df <- dplyr::ungroup(df)

df <- dplyr::mutate(df,
  height_cm = ifelse(is.na(height_cm), mean(height_cm, na.rm = TRUE), height_cm),
  weight_kg = ifelse(is.na(weight_kg), mean(weight_kg, na.rm = TRUE), weight_kg)
)
print(df)
```
Write your function(s) in the cell below:
```{r}
```
## Exercise 2: Add your functions to the pipeline

**Tasks:**
- Add your functions to an appropriate module in the `R/` folder
- Ensure your functions are used when running the pipeline

**Hints:**
- Add your functions into another function in the same module
- If you create a new module, import the module into the main.R script

## Exercise 3: Challenge & Reflection

**Tasks:**
- Compare the cleaned_data output with and without your changes
- Document your modules with Roxygen documentation

**Reflect:**
- How does modular code improve reproducibility and maintainability?
- What are the benefits of separating code into modules?
- How would you extend this pipeline to add new features or analyses?

**Bonus:**
- Try adding your own modules and functions to the pipeline

## Exercise 4: Refactor Visualisation Code

As an additional challenge, Add code to create visualisations. Place all
visualisation functions in the `report.R` module.

**Tasks:**
- Write a function to plot missing values per column before the data is cleaned
- Write a function to plot disease prevalence for each disease category over time
after the data is cleaned
- Add the visualisations to the output report

**Bonus:**
- Save the charts to the reports folder
- Customise chart formatting and colours
