# Unit Testing Exercises

Unit tests help ensure your code works as expected and make your RAP
pipeline more robust and reproducible. In this exercise, you'll learn
how to write and run unit tests.

## Why are unit tests important?

Unit tests check that individual functions behave as expected. They: -
- Help catch bugs, by testing small pieces of code in isolation
- Make code easier to maintain by documenting expected behavior
- Ensure code changes don't break existing functionality by running tests after
modifications
- Give confidence when refactoring or adding new features by verifying existing
tests still pass

Read more in the [testthat documentation](https://testthat.r-lib.org/)

```{r setup, include = FALSE}
# Setup code for exercises

library(dplyr)

input_path <- here::here("data", "inputs", "health_data.csv")
```

## Exercise 1: Review an existing unit test

The function `clean_health_data` in `R/cleaning.R` has an existing unit test in
`tests/test_cleaning.R`.

**Task:** 1. Run the existing unit test for `clean_health_data` to understand 
how it works. 2. Modify the sample DataFrame in the test to understand what
causes the tests to pass or fail (Note: if all rows are dropped all tests will fail).

## Exercise 2: Write a simple unit test for a new function

Suppose you have created a function called `impute_by_group` in
`R/cleaning.R`.

**Task:** 1. Create a new test file in the `tests/` folder (e.g.,
`test_cleaning.R`). 2. Write a unit test for `impute_by_group` that
checks it correctly imputes missing values. 3. Use testthat for your
test.

**Hints:** - Read on for more guidance

## Exercise 2a: Walkthrough - Write a unit test for flag_missing

Let's start with a simple function called `flag_missing`. This function
adds a new column to your DataFrame to flag missing values in specified
columns.

Below is the code for `flag_missing`:

```{r}
library(dplyr)

flag_missing <- function(df) {
  df <- df %>%
    dplyr::rowwise() %>%
    dplyr::mutate(height_cm_imputed = is.na(height_cm)) %>%
    dplyr::mutate(weight_kg_imputed = is.na(weight_kg)) %>%
    dplyr::ungroup()
  return(df)
}

# Example output
df <- data.frame(
  height_cm = c(170, NA, 160),
  weight_kg = c(70, 80, NA)
)
flagged <- flag_missing(df)
```

## How to write a unit test for `flag_missing`

1.  Add the `flag_missing` function to the `R/cleaning.R`
    file if you have not done so already.
2.  Open the `test_cleaning.R file` in the `tests/` folder.
3.  The source function in `test_cleaning.R` will automatically import
    all of the functions saved in the `cleaning.R` file.
4.  Write a test function using `test_that`:

```{r}
library(testthat)
# If running the test in a separate file to the function, use source() to import
# the function that is being tested

testthat::test_that("test flag_missing", {
  # Create data frame to feed into the flag_missing function
  df <- data.frame(
    height_cm = c(170, NA, 160),
    weight_kg = c(70, 80, NA)
  )
  # Define a data frame of the expected output that is expected from flag_missing
  expected_output <- data.frame(
    height_cm = c(170, NA, 160),
    weight_kg = c(70, 80, NA),
    height_cm_imputed = c(FALSE, TRUE, FALSE),
    weight_kg_imputed = c(FALSE, FALSE, TRUE)
  )

  flagged <- flag_missing(df)
  # Check that the output of flag_missing is the same as the expected output
  testthat::expect_equal(as.data.frame(flagged), expected_output)
})
```

5.  Run the tests by running devtools::test()
6.  If the test passes, your function works as expected. If it fails,
    check the error message and review your code.

#### Understanding the test_flag_missing function

Let's break down how the `test_flag_missing` function works:

-   **Creating test data:** We create a small DataFrame `df` with some
    missing values. This lets us check if our function works as
    expected.
-   **Calling the function:** We call `flag_missing` and pass in our
    test DataFrame.
-   **Checking the results:** We use and `expect` statement such as an
    `expect_equal` statement to check that the new columns (e.g.,
    `height_cm_imputed`, `weight_kg_imputed`) correctly show `True` for
    missing values and `False` otherwise. If the output matches what we
    expect, the test will pass.
-   **Why use expect statements?** Expect statements are a way to say
    "this must be true". If the condition is not true, testthat will
    show an error and the test will fail.

This approach helps you check that your function does what you expect,
using simple, readable code. You can add more assert statements or test
with different data to make your tests even stronger.

## Exercise 2b: Write a unit test for impute_by_group

Now try writing your own tests for the following function. Use the
walkthrough above as a guide.

**Hints:** - Import the function at the top of your test file. - Use a
small DataFrame with missing values for testing. - Use `expect`
statements to check the output.

Below is the function definition:

```{r}
library(dplyr)

impute_by_group <- function(df, col, group_col) {
  col <- rlang::sym(col)
  group_col <- rlang::sym(group_col)

  df <- df %>%
    dplyr::group_by(!!group_col) %>%
    dplyr::mutate(
      imputed = ifelse(
        is.na(!!col),
        mean(!!col, na.rm = TRUE),
        !!col
      )
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      imputed = ifelse(
        is.na(imputed),
        mean(imputed, na.rm = TRUE),
        imputed
      ),
      imputed = round(imputed, 1)
    )

  return(df$imputed)
}

# Example output
df <- read.csv(input_path)
impute_by_group(df = df, col = "weight_kg", group_col = "gender")
```

**Task:** - Add the `flag_missing`, `impute_by_group` functions to the
`R/cleaning.R` file if you have not done so already. 
- Open the `test_cleaning.R` file in the `tests/` folder. 
- Write a unit test for `impute_by_group` that checks missing values
are correctly imputed using group means. 
- Use testthat and expect statements in your test file.

**Note:** The tests will run for these functions, but the functions will not
affect the output unless you implement them in the pipeline. See
`02_modules_solutions` for a guide on how to do this

## Exercise 3: Run your unit tests

**Task:** - To run tests in the `tests/` folder you must open each test
one at a time and click the `run tests` button in the top right corner
of the source pane or use devtools::test(). 
- Check the output to see if your tests pass or fail.

## Exercise 4: Stretch - Check test coverage

Test coverage shows how much of your code is tested by unit tests.

**Task:** - Import covr: `library(covr)`

- Run: `file_coverage(path_to_function, path_to_test)` to get a
percentage of the lines of code that are covered by the test.

To which lines of code are not covered by the test, set a variable (for
example: cov) equal to the code above and then run:

`zero_lines <- covr::zero_coverage(cov)`

`print(zero_lines)` - Check which lines of your code are covered by
tests.

## Exercise 5: Stretch - Try parameterisation in testthat

Parameterisation lets you run the same test with different inputs.

**Task:** - Install the R package patrick - Use
`with_parameters_test_that` in your test file to test your function with
multiple input cases. - See the [patrick
documentation](https://github.com/google/patrick/blob/main/README.md)
for examples.
